On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-11 Revisiones/PNA_MMP_Responses to the reviewers.tex

Of course, the employment of the \textit{max-ent} approximants is not free of difficulties. Some requirements such as the need for a convex domain, as well as those derived from its calculation (the determination of the Lagrange Multipliers, specially at the boundaries, or the obtainment of the minimizer of the logarithmic function), make the usage of it a challenging tool. However, the results depicted in this manuscript will strengthen the motivation of its employment in order to mitigate typical MPM problems such as cell-crossing, stress instabilities and the use of unstructured mesh.

Concerning the first issue mentioned by the reviewer, the special case of non-convex domains was discussed by Arroyo \& Ortiz~\cite{Arroyo2006}. Some of the solutions proposed by the aforementioned authors are : the possibility of replacing the Euclidean distance $\lVert  x - x_a  \rVert$ in the definition of the shape functions by the length of the shortest path contained within the domain connecting $x$ and $x_a$. Or the decomposition of the non-convex domain into convex sub-domains.  This topic has also been extensively studied in the context of MLS-based meshfree methods, for instance visibility, diffraction, and constrained path criteria. These methods are directly applicable to local \textit{max-ent} approximation.

\item \textit{Does the use of max-ent solve the problem of imposing Dirichlet boundary conditions in the MPM?}\\

Unfortunately, local \textit{max-ent} approximation technique only provides an efficient tool to transfer information from particles to nodes and vice versa \sout{to the MPM}. Hence, the difficulties of imposing Dirichlet boundary conditions (DBC) in the MPM persists since they are a direct consequence of discretising the continuum with material points instead of nodes, where traditionally DBC are imposed. However, any available technique employed to overcome this situation in the standard MPM can be translated to the \textit{max-ent} one thanks to the Kronecker-Delta property inherent to it \cite{Arroyo2006}. See for instance the work of Cortis {\it et al.} (2018) \cite{Cortis_et_al_2017_IJNME} as example of how DBC can be imposed in the MPM.

@@

These interesting publications will be added to the state of art of the revised manuscript. In the humble opinion of the authors, the XPIC approach is dedicated to mitigate the null space error produced during the transfer of information between particles and nodes. Therefore, it is related to the spatial discretisation instead of the temporal one. Meanwhile, the algorithm introduced in this research proposes the stabilisation within the adaptation of the predictor-corrector algorithm, which has been used traditionally in FEM to deal with fast dynamics. To the author\textquotesingle s knowledge, this is the first attempt to adapt the  traditional predictor-corrector scheme for the MPM framework.

\end{enumerate}

 \textit{The numerical examples are not impressive when the message of the paper is that this is good for fast dynamics problems. Compare to papers such as Ma et al. (2009) IJ Impact Engineering}\\

\sout{Contrary to the reviewer\textquotesingle s opinion,} In the authors' opinnion, the provided results are capable to highlight the \sout{strong} capabilities of the \acrshort{LME-}MPM-NPC proposed in the present work. It is important to remark the extension of this research: in order to isolate the strength of the proposed methodology, covering spatial and time discretization, only linear elastic problems are assessed.

### 2020-11-09 LME-MPM.tex
Updates from Overleaf
Version para el CMAME

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. \textcolor{blue}{The described methodology is sketched by the scheme of the figure~\ref{fig:MPM_algorithm}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme_horizontal}
  }
  \caption{Description of the three steps in \acrshort{mpm} standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In what follows, we will adopt the following convention: \textcolor{blue}{Any vector field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are  considered:  subscript $\square_p$ is used to define a particle variable,  subscript $\square_I$ is reserved for nodal variables, superscript $\square^{\psi}$ involves a virtual magnitude. \textcolor{blue}{The convention adopted for the operators is as folows:} $\dot{\square}$ and $\ddot{\square}$ \textcolor{blue}{are considered} for the first and second time derivative, $\square \otimes \square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ \textcolor{blue}{are the single and double contraction of tensor index}, $\Div{\square}$ denotes the divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{red}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$  \textcolor{red}{where} $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{blue}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
\textcolor{blue}{together with essential or \acrfull{dbc} at \gls{dirichlet-boundary}.} We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{red}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
\textcolor{blue}{together with the natural of \acrfull{nbc} at \gls{neumann-boundary}. An additional component is (iii) the constitutive equation which relates stress and strain increments as,}
{\color{red}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\Delta \sigma} = \tens{D} \colon \tens{\Delta \varepsilon}.
\end{equation}
}
\textcolor{blue}{
In this research, plane strain linear elasticity is considered. The final equation of the set is (iv) the mass conservation, which can be expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

@@

    \includegraphics[width=\textwidth]{Figures/Mesh-particles-back}}

@@

@@

obtained, the \acrshort{dbc} are imposed over \gls{dirichlet-boundary}. \textcolor{blue}{Next the \textit{corrector}  step is introduced. Predicted nodal velocities obtained in \eqref{eq:Predictor-velocity-II}, are now corrected by the equation}

@@

      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/MPM_Shape_Fun_dy}

@@

      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_17_3_Shape_Fun_dy}

@@

      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/GIMP_Shape_Fun_dy}

@@

    \begin{tabular}{c} \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun}\\ \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun_dx}\\      \includegraphics[width=0.14\textwidth]{Figures/LME_10_0_Shape_Fun_dy}    

@@

      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun_dx}\\
      \includegraphics[width=0.14\textwidth]{Figures/LME_5_0_Shape_Fun_dy}

@@

This benchmark was proposed by Dyka~\cite{Dyka1995} since allows to study easily the capability
of the proposed time integration algorithm to avoid velocity
field instabilities. It consists of a one-dimensional bar of a length
of 0.1333 meters, sketched in the figure~\ref{fig:Dyka_Bar}. The
boundary conditions are: displacements are constrained ($\vec{v}
\rvert_{x=L} = 0$) in the right border, being free on all other boundaries. An initial velocity of $\vec{v}_o = - 5\ m/s$ is given to the
left quarter of the bar. Finally, the elastic parameters chosen for this test are:
\begin{itemize} 
\item  Density : $7833\ kg/m3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $200 \cdot 10^9\ Pa$
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Dyka_bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the proposed example, the simulation extends from time zero up to \textcolor{red}{$10^{-4}$} s.
This time interval allows the elastic wave to travel a distance of 2.6
lengths of the bar. For the spatial discretisation, a set of seven nodal mesh sizes (0.1, 0.3325, 0.5, 1.0, 3.3325, 6.665, 10.0 millimeters) are considered. For each element a number of four particles was selected. In the initial layout, particles are located the exact quadrature points of a linear quadrilateral, with the exception of the \acrshort{ugimp} simulation, where gaps or overlap between
voxels of each particle are not allowed. In those cases, each particle occupies the center of each cell quarter. For all simulations, time step is controlled by a Courant-Friedrichs-Levy condition of 0.1 , were the adopted celerity is computed as:
\begin{equation}
  \label{eq:Cel}
  Cel = \max\{\max_{p \in \Omega_p}\{ \vec{v}_p \} , \max_{p \in \Omega_p}\{ \sqrt{\frac{E_p}{\rho_p}} \} \}.
\end{equation}
An important consideration regarding modellization concerns the
background mesh. Notice that free border of the bar has a maximum
horizontal displacement of 0.03 millimeters, therefore 
a computational domain with an extra gap of 0.03 millimeters is
required in order to accommodate the unconstrained displacement of the
particles in the left border of the bar. This problem arises
when the mesh size is small enough that relative displacement of the particles is larger \textcolor{red}{than} the distance to the border, so grid crossing phenomena could appear even in those cases with infinitesimal displacements. In this case, an analytical solution can be obtained through the characteristics method, described in the appendix~\ref{app:analytical_sol}. \textcolor{blue}{This section contains a large number of comparisons, which have been summarised in Table \ref{tab:dyka-cases-summarized}. First, a convergence study regarding the integration scheme and interpolation technique is performed in  \ref{sec:Convergence-analysis}. Second, comparison of the capabilities of the \acrshort{npc} and \acrshort{fe} algorithms to damp out numerical noise is analysed in \ref{sec:FE-vs-NPC}. Third, the response of the \acrshort{lme} approximation for different $\widehat{\gamma}$ values is study in  \ref{sec:Sensitive-analysis-gamma}. Fourth, comparison between the well-known \acrshort{ugimp} shape function against the \acrshort{lme} approximation is performed in \ref{sec:LME-vs-uGIMP}. Finally, a comparison between the \acrshort{mpm} and the \acrfull{otm} method, using the same time integration scheme and interpolation technique, is conducted in \ref{sec:OTM-vs-MPM}.}

@@

To measure the convergence of the solutions
for the different time integration and approximation schemes the
root-mean-square (RMS) error in the velocity field is computed. RMS
error is defined as
\begin{equation}
  \label{eq:RMS}
  RMS = \sqrt{\frac{1}{N} \sum^{N}_p \left( \vec{v}_p - \hat{\vec{v}}_p \right)^2},
\end{equation}
where $\vec{v}_p$ and $\hat{\vec{v}}_p$ are respectively the analytical and
numerical solutions evaluated in the final time step in the position
of each particle. In Fig. \ref{fig:Dyka-error-evol} the evolution of the RMS is obtained for both time integration schemes. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Error_evol}
  }
  \caption{Velocity error evolution at the point A in the Dyka's bar ,
    convergence plots for \acrshort{fe} and \acrshort{npc}. The plot is subdivided with
    colours, the darker part of the diagram shows coincides when the
    relative movement of the particles is large enough to produce the
    grid crossing phenomena. The lightest part of the diagram
    coincides when the relative movement of the particles in
    negligible in comparison with the mesh size. And in the middle
    region a transition behaviour take place.}
  \label{fig:Dyka-error-evol}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The right figure, with the \acrshort{npc} results, shows lower values of the estimated error, denoting the higher performance of this methodology. About the spatial discretisation, the \acrshort{lme} schemes show an error comparable to the obtained with the \acrshort{ugimp}, being even lower close to the \textit{grid-crossing region}. \textcolor{blue}{In this region the performance is punished with significant movement of the particles as far as the mesh size is reduced.}

@@

\textcolor{blue}{A comparison of the performance of both time integration schemes is presented in figure \ref{fig:Dyka-NPC-FE}. Both integration schemes show numerical oscillations with respect to the analytical solution all over the simulation time, being worse just after jump discontinuities. However, the \acrshort{npc} approach damps out these oscillations faster than the \acrshort{fe} scheme. The oscillations of the \acrshort{npc} almost disappeared before reaching subsequent jump discontinuities, while those of \acrshort{fe} do not. Moreover, oscillations of the \acrshort{fe} seems to increase further after each jump discontinuity and an unstable tendency can by foreseen. This unstable tendency is not appreciated in the \acrshort{npc} integration scheme.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{Comparison of \acrshort{npc} and \acrshort{fe}
      performances: In the picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\textcolor{blue}{Figure \eqref{fig:Dyka-LME-gamma} shows the sensitivity of the \acrshort{lme}
approximation scheme to variations in the parameter
$\widehat{\gamma}$, which is the one that controls the value of the regularization parameter
$\beta$ together with the nodal spacing parameter $h$. It can be observed that for lower values of $\widehat{\gamma}$ the numerical solution presents a faster decay of the spurious oscillations. This capability could be useful in simulations where
extremely noisy oscillations could damage the solutions like memory materials.  An additional observation, concerning to the
solution sensibility depending on the regularization parameters, is the behaviour of the solution depending on the decreasing of mesh size. For larger mesh sizes, where the relative particle displacement is negligible in comparison with the cell size, the global behaviour is \acrshort{fem}-like, therefore, larger values of $\widehat{\gamma}$ may offer better results. On the other hand, when mesh size is small enough to produce grid-crossing, meshfree behaviour is required to ensure the convergence of the solution and tiny values of $\widehat{\gamma}$ may lead to better performances.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_LME_gamma_comparative}
  }
  \caption{Sensitivity  of LME approximants performance to changes in the
    dimensionless regularization parameter \textcolor{blue}{$\widehat{\gamma}$}. To illustrate it, the velocity evolution at the point in the bar left side is plotted.}
  \label{fig:Dyka-LME-gamma}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% name your BibTeX data base
\bibliography{Biblio/Biblio}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method consists of three main steps: (i) a
variational recovery process, where particle data is projected onto the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. \textcolor{blue}{The described methodology is sketched by the scheme of the figure~\ref{fig:MPM_algorithm}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures-MPM-scheme-horizontal}
  }
  \caption{Description of the three steps in \acrshort{mpm} standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 In what follows, we will adopt the following convention: \textcolor{blue}{Any vector field is defined with an overarrow as $\vec{\square}$}. Three kind of subscripts or superscripts are  considered:  subscript $\square_p$ is used to define a particle variable,  subscript $\square_I$ is reserved for nodal variables, superscript $\square^{\psi}$ involves a virtual magnitude. \textcolor{blue}{The convention adopted for the operators is as folows:} $\dot{\square}$ and $\ddot{\square}$ \textcolor{blue}{are considered} for the first and second time derivative, $\square \otimes \square$ means the dyadic operator, $\square \cdot \square$ and $\square \colon \square$ \textcolor{blue}{are the single and double contraction of tensor index}, $\Div{\square}$ denotes the divergence operator, and finally $\Grad{\square}$ and $\GradS{\square}$ denotes the gradient and its symmetric part. Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are \textcolor{red}{described} in subsection \ref{sec:derivation-mpm}, \ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.

@@

In \acrshort{mpm} the continuum mechanics approach is considered. We consider a region \gls{domain} occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \Omega
= \Gamma_d \cup \Gamma_n$  \textcolor{red}{where} $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures-Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \Omega$
    of the domain $\Omega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. First (i) we relate global to local state by the \textit{compatibility equation} \textcolor{blue}{where the strain field \gls{strain} is defined by}:
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vec{u}},
\end{equation}
\textcolor{blue}{together with essential or \acrfull{dbc} at \gls{dirichlet-boundary}.} We will further assume that 
strains are infinitesimal. The stress field \gls{stress} is considered the corresponding
conjugate variable for the strain field, being the one which
satisfies (ii) the \textit{conservation of \textcolor{red}{linear} momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vec{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vec{b}
\end{equation}
\textcolor{blue}{together with the natural of \acrfull{nbc} at \gls{neumann-boundary}. An additional component is (iii) the constitutive equation which relates stress and strain increments as,}
{\color{red}
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\Delta \sigma} = \tens{D} \colon \tens{\Delta \varepsilon}.
\end{equation}
}
\textcolor{blue}{
In this research, plane strain linear elasticity is considered. The final equation of the set is (iv) the mass conservation, which can be expressed by} 
\begin{equation}
  \label{eq:Rho-material-derivative}
    \dot{\rho} + \rho \Div{\vec{v}} = 0.
\end{equation}

@@

    \includegraphics[width=\textwidth]{Figures-Mesh-particles-back}}

@@

  \label{eq:Predictor-velocity-III}

@@

obtained, the \acrshort{dbc} are imposed over \gls{dirichlet-boundary}. \textcolor{blue}{Next the \textit{corrector}  step is introduced. Predicted nodal velocities obtained in \eqref{eq:Predictor-velocity-III}, are now corrected by the equation}

@@

      \includegraphics[width=0.14\textwidth]{Figures-MPM-Shape-Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures-MPM-Shape-Fun-dx}\\
      \includegraphics[width=0.14\textwidth]{Figures-MPM-Shape-Fun-dy}

@@

      \includegraphics[width=0.14\textwidth]{Figures-LME-17-3-Shape-Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures-LME-17-3-Shape-Fun-dx}\\
      \includegraphics[width=0.14\textwidth]{Figures-LME-17-3-Shape-Fun-dy}

@@

      \includegraphics[width=0.14\textwidth]{Figures-GIMP-Shape-Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures-GIMP-Shape-Fun-dx}\\
      \includegraphics[width=0.14\textwidth]{Figures-GIMP-Shape-Fun-dy}

@@

     \includegraphics[width=0.14\textwidth]{Figures-LME-10-0-Shape-Fun}\\ 		   \includegraphics[width=0.14\textwidth]{Figures-LME-10-0-Shape-Fun-dx}\\      \includegraphics[width=0.14\textwidth]{Figures-LME-10-0-Shape-Fun-dy}    

@@

      \includegraphics[width=0.14\textwidth]{Figures-LME-5-0-Shape-Fun}\\
      \includegraphics[width=0.14\textwidth]{Figures-LME-5-0-Shape-Fun-dx}\\
      \includegraphics[width=0.14\textwidth]{Figures-LME-5-0-Shape-Fun-dy}

@@

This benchmark was proposed by Dyka~\cite{Dyka1995} since allows to study easily the capability
of the proposed time integration algorithm to avoid velocity
field instabilities. It consists of a one-dimensional bar of a length
of 0.1333 meters, sketched in the figure~\ref{fig:Dyka_Bar}. The
boundary conditions are: displacements are constrained ($\vec{v}
\rvert_{x=L} = 0$) in the right border, being free on all other boundaries. An initial velocity of $\vec{v}_o = - 5\ m/s$ is given to the
left quarter of the bar. Finally, the elastic parameters chosen for this test are:
\begin{itemize} 
\item  Density : $7833\ kg/m3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $200 \cdot 10^9\ Pa$
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures-Dyka-bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the proposed example, the simulation extends from time zero up to \textcolor{red}{$10^{-4}$} s.
This time interval allows the elastic wave to travel a distance of 2.6
lengths of the bar. For the spatial discretisation, a set of seven nodal mesh sizes (0.1, 0.3325, 0.5, 1.0, 3.3325, 6.665, 10.0 millimeters) are considered. For each element a number of four particles was selected. In the initial layout, particles are located the exact quadrature points of a linear quadrilateral, with the exception of the \acrshort{ugimp} simulation, where gaps or overlap between
voxels of each particle are not allowed. In those cases, each particle occupies the center of each cell quarter. For all simulations, time step is controlled by a Courant-Friedrichs-Levy condition of 0.1 , were the adopted celerity is computed as:
\begin{equation}
  \label{eq:Cel}
  Cel = \max\{\max_{p \in \Omega_p}\{ \vec{v}_p \} , \max_{p \in \Omega_p}\{ \sqrt{\frac{E_p}{\rho_p}} \} \}.
\end{equation}
An important consideration regarding modellization concerns the
background mesh. Notice that free border of the bar has a maximum
horizontal displacement of 0.03 millimeters, therefore 
a computational domain with an extra gap of 0.03 millimeters is
required in order to accommodate the unconstrained displacement of the
particles in the left border of the bar. This problem arises
when the mesh size is small enough that relative displacement of the particles is larger \textcolor{red}{than} the distance to the border, so grid crossing phenomena could appear even in those cases with infinitesimal displacements. In this case, an analytical solution can be obtained through the characteristics method, described in the appendix~\ref{app:analytical_sol}. \textcolor{blue}{This section contains a large number of comparisons, which have been summarised in Table \ref{tab:dyka-cases-summarized}. First, a convergence study regarding the integration scheme and interpolation technique is performed in  \ref{sec:Convergence-analysis}. Second, comparison of the capabilities of the \acrshort{npc} and \acrshort{fe} algorithms to damp out numerical noise is analysed in \ref{sec:FE-vs-NPC}. Third, the response of the \acrshort{lme} approximation for different $\widehat{\gamma}$ values is study in  \ref{sec:Sensitive-analysis-gamma}. Fourth, comparison between the well-known \acrshort{ugimp} shape function against the \acrshort{lme} approximation is performed in \ref{sec:LME-vs-uGIMP}. Finally, a comparison between the \acrshort{mpm} and the \acrfull{otm} method, using the same time integration scheme and interpolation technique, is conducted in \ref{sec:OTM-vs-MPM}.}

@@

To measure the convergence of the solutions
for the different time integration and approximation schemes the
root-mean-square (RMS) error in the velocity field is computed. RMS
error is defined as
\begin{equation}
  \label{eq:RMS}
  RMS = \sqrt{\frac{1}{N} \sum^{N}_p \left( \vec{v}_p - \hat{\vec{v}}_p \right)^2},
\end{equation}
where $\vec{v}_p$ and $\hat{\vec{v}}_p$ are respectively the analytical and
numerical solutions evaluated in the final time step in the position
of each particle. In Fig. \ref{fig:Dyka-error-evol} the evolution of the RMS is obtained for both time integration schemes. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure*}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures-Error-evol}
  }
  \caption{Velocity error evolution at the point A in the Dyka's bar ,
    convergence plots for \acrshort{fe} and \acrshort{npc}. The plot is subdivided with
    colours, the darker part of the diagram shows coincides when the
    relative movement of the particles is large enough to produce the
    grid crossing phenomena. The lightest part of the diagram
    coincides when the relative movement of the particles in
    negligible in comparison with the mesh size. And in the middle
    region a transition behaviour take place.}
  \label{fig:Dyka-error-evol}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The right figure, with the \acrshort{npc} results, shows lower values of the estimated error, denoting the higher performance of this methodology. About the spatial discretisation, the \acrshort{lme} schemes show an error comparable to the obtained with the \acrshort{ugimp}, being even lower close to the \textit{grid-crossing region}. \textcolor{blue}{In this region the performance is punished with significant movement of the particles as far as the mesh size is reduced.}

@@

\textcolor{blue}{A comparison of the performance of both time integration schemes is presented in figure \ref{fig:Dyka-NPC-FE}. Both integration schemes show numerical oscillations with respect to the analytical solution all over the simulation time, being worse just after jump discontinuities. However, the \acrshort{npc} approach damps out these oscillations faster than the \acrshort{fe} scheme. The oscillations of the \acrshort{npc} almost disappeared before reaching subsequent jump discontinuities, while those of \acrshort{fe} do not. Moreover, oscillations of the \acrshort{fe} seems to increase further after each jump discontinuity and an unstable tendency can by foreseen. This unstable tendency is not appreciated in the \acrshort{npc} integration scheme.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures-Velocity-FE-vs-PCE-CFL-01}
  }
  \caption{Comparison of \acrshort{npc} and \acrshort{fe}
      performances: In the picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\textcolor{blue}{Figure \eqref{fig:Dyka-LME-gamma} shows the sensitivity of the \acrshort{lme}
approximation scheme to variations in the parameter
$\widehat{\gamma}$, which is the one that controls the value of the regularization parameter
$\beta$ together with the nodal spacing parameter $h$. It can be observed that for lower values of $\widehat{\gamma}$ the numerical solution presents a faster decay of the spurious oscillations. This capability could be useful in simulations where
extremely noisy oscillations could damage the solutions like memory materials.  An additional observation, concerning to the
solution sensibility depending on the regularization parameters, is the behaviour of the solution depending on the decreasing of mesh size. For larger mesh sizes, where the relative particle displacement is negligible in comparison with the cell size, the global behaviour is \acrshort{fem}-like, therefore, larger values of $\widehat{\gamma}$ may offer better results. On the other hand, when mesh size is small enough to produce grid-crossing, meshfree behaviour is required to ensure the convergence of the solution and tiny values of $\widehat{\gamma}$ may lead to better performances.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures-Velocity-LME-gamma-comparative}
  }
  \caption{Sensitivity  of LME approximants performance to changes in the
    dimensionless regularization parameter \textcolor{blue}{$\widehat{\gamma}$}. To illustrate it, the velocity evolution at the point in the bar left side is plotted.}
  \label{fig:Dyka-LME-gamma}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% name your BibTeX data base
\bibliography{Biblio}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
